Foundations for Inference

### 2018-10-16 DATA606/Foundations for Inference.Rmd
Updated DATA 606 Homework

---
title: "Foundations for Inference"
author: "Mike Silva"
date: "October 16, 2018"
output: html_document
---

@@

(a) What is the point estimate for the average height of active individuals? What about the median?

(b) What is the point estimate for the standard deviation of the heights of active individuals? What about the IQR?

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning.

(d) The researchers take another random sample of physically active individuals. Would you expect the mean and the standard deviation of this new sample to be the ones given above? Explain your reasoning.

(e) The sample means obtained are point estimates for the mean height of all active individuals, if the sample of individuals is equivalent to a simple random sample. What measure do we use to quantify the variability of such an estimate (Hint: recall that $SD_\bar{x} = \frac{\sigma}{\sqrt{n}}$)? Compute this quantity using the data from the original sample under the condition that the data are a simple random sample.

@@

(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11.

(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.

(c) 95% of random samples have a sample mean between $80.31 and $89.11.

(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11.

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don’t need to be as sure about our estimate.

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger.

(g) The margin of error is 4.4.

@@

(a) Are conditions for inference satisfied?

@@

Define the term “sampling distribution” of the mean, and describe how the shape, center, and spread of the sampling distribution of the mean change as sample size increases.

@@

(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours?

(b) Describe the distribution of the mean lifespan of 15 light bulbs.

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?

@@

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution?

>>>>

---
title: "Foundations for Inference"
author: "Mike Silva"
date: "October 16, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---

@@

(a) What is the point estimate for the average height of active individuals? **171.1 cm** What about the median? **170.3 cm**

(b) What is the point estimate for the standard deviation of the heights of active individuals? **9.4 cm** What about the IQR? **177.8 cm - 163.8 cm = 14 cm**

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning.  **A person who is 180 cm is pretty tall but I wouldn't call it unusually tall b/c it's within 1 standard deviation of the mean.  By the same token, I would consider someone who is 155 cm small because they are less than one standard deviation from the mean.**

(d) The researchers take another random sample of physically active individuals. Would you expect the mean and the standard deviation of this new sample to be the ones given above? Explain your reasoning. **I would expect both the mean and standard deviation to be different.  There is variation within the population which would translate into a different sample mean and standard deviations.**

(e) The sample means obtained are point estimates for the mean height of all active individuals, if the sample of individuals is equivalent to a simple random sample. What measure do we use to quantify the variability of such an estimate (Hint: recall that $SD_\bar{x} = \frac{\sigma}{\sqrt{n}}$)? **The standard error**  Compute this quantity using the data from the original sample under the condition that the data are a simple random sample. **0.4174687**
$SD_\bar{x} = \frac{\sigma}{\sqrt{n}} = \frac{9.4}{\sqrt{507}} = 0.4174687$

@@

(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11. **False.  The 95% applies to the population parameter not the sample.**

(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.  **True.  This is one of the conditions for a confidence interval.**

(c) 95% of random samples have a sample mean between $80.31 and $89.11. **False.  Cofidence interval does not contstrain other sample means.**

(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11. **True.  This is assuming that the 436 people in the sample were a renadom draw.**

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don’t need to be as sure about our estimate. **False.  We need to be just as sure about the estimate.  The 90% means that the population parameter is within the CI 90% of the time or that we are less sure that our sample captures the population parameter.**

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger. **False.  If we had a large sample size trippling it would not decrease the margin of error.**

(g) The margin of error is 4.4. **False.  That is the difference between the lower bound of the CI and the mean, not the margin of error.**

@@

(a) Are conditions for inference satisfied? **The N > 30 rule of thumb is satisfied, the children were randomly selected, and the histogram appears to be normally distributed.  I think all conditions have been met.**

@@

Define the term “sampling distribution” of the mean, and describe how the shape, center, and spread of the sampling distribution of the mean change as sample size increases.  **This is the distributions of the mean value of the samples.  It is normally distributed.  The mean approximates the population mean.  As the sample size increases the sample mean approaches the population mean, and the spread decreases.**

@@

(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours? **0.067**

```{r}
mu <- 9000 
sd <- 1000
x <- 10500
1 - pnorm(x, mu, sd)
```

(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**

@@

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**
